# Literature Review

```{r setup, file = "R/chapter_start.R", include = FALSE}
# a number of commands need to run at the beginning of each chapter. This
# includes loading libraries that I always use, as well as options for 
# displaying numbers and text.
```

At the end of the 1990s, the Kansas DOT sought an alternative to temporary cold-mix asphalt rumble strips it used to alert drivers to upcoming work zones [@meyerEvaluationOrangeRemovable2000]. About ten years of product development led from adhesive strips pasted to the roadway to non-adhesive steel and rubber pads heavy enough to resist movement but light enough to be handled by workers [@maze2000; @meyer2006; @brown22].

Meyer et al. (2006) developed specifications for a non-adhesive rumble strip addressing limitations identified for adhesive strips. In December 2008, Plastic Safety Systems (PSS) released the RoadQuake, a non-adhesive strip 3/4 in. thick, made of rubber and steel (PSS, 2018). Studies using both track and field tests demonstrated that RoadQuake and similar non-adhesive TPRS generated more noise feedback than adhesive TPRS and remained in place when properly installed (Schrock et al., 2010; Schrock et al., 2016; Sun et al., 2011). Research on adhesive TPRS since the development of non-adhesive TPRS is lacking, indicating a strong industry shift. Unless stated otherwise, all mentions of TPRS in the remainder of this report refer to the non-adhesive variety.

## Overview

Non-adhesive TPRS, typically made of steel and rubber, have been used since 2009 to improve work zone safety [@meyer_design_2006, @pss_best_2018]. Research consistently shows that TPRS can reduce traffic speed, increase driver attentiveness, and minimize erratic driving behaviors such as lane departures [@brown_effectiveness_2022, @sun_elevated-risk_2011, @wang_effects_2013, @yang_effectiveness_2015].

## Speed Reduction

TPRS installations are effective at reducing vehicle speeds in work zones. Studies report speed reductions ranging from 3.5 to 13.8 mph,[@brown_effectiveness_2022, @sun_elevated-risk_2011, @wang_effects_2013, @yang_effectiveness_2015], with greater reductions observed near the first set of strips and when drivers can see a flagger or additional barriers [@wang_effects_2013]. @wang_effects_2013 shows the effect is highly localized, and speeds tend to increase once drivers pass the TPRS unless additional measures are present. 


## Driver Behavior

TPRS increase driver attentiveness, often measured by the frequency of brake light activation as vehicles approach the strips. Studies show that TPRS prompt more drivers to brake and pay attention to changing road conditions, though the percentage varies from 11 perecnt to 55 percent based on site and conditions [@brown_effectiveness_2022; @sun_elevated-risk_2011; @wang_effects_2013; @yang_effectiveness_2015]. Commercial vehicles tend to respond more cautiously than passenger cars [@wang_effects_2013].

Concerns about drivers evading TPRS by leaving their lane have been largely unfounded. Most studies report minimal lane departures, especially when TPRS are installed across the full width of the road [@sun_elevated-risk_2011; @wang_effects_2013; @brown_effectiveness_2022]. Lane departures are more likely on narrow roads or near the end of a TPRS array, but overall rates are low [@sun_elevated-risk_2011].

## TPRS Displacement

A key challenge with non-adhesive TPRS is their tendency to shift or become displaced under traffic. Displacement increases with heavier vehicles, higher speeds, and higher traffic volume [@heaslip_closed-course_2010, @schrock_development_2016, @sun_elevated-risk_2011]. Placement angle also matters—strips that rotate away from perpendicular to traffic tend to move more[@sun_elevated-risk_2011]. Real-world studies show that while displacement is generally manageable, it can require periodic adjustment, which may increase worker exposure to traffic. The effect of strip wear and long-term durability under heavy use is not well understood and requires further research.

## Spacing Specifications

There are three main groups of spacing specifications. Some states, such as Wisconsin, Florida, and Colorado, specify a single spacing independent of speed [e.g., @wdotSDD15C12]. Some states such as Maryland, Viriginia, Minnesota, and Utah vary spacing from 10 to 20 ft. based on the posted speed limit [@minndot2024]. New York, Texas, and Missouri specify spacing from 10 to 35 ft. based on the posted speed limit[@citation]. The manufacturer also recommends spacing of 10 to 35 ft. based on the posted speed limit[@citation]. The various specifications are shown in Figure @fig-state-specs.


```{r}
#| label: fig-state-specs
#| fig-cap: "Various specifications for TPRS spacing."
tar_load(state_spacing_plot)
state_spacing_plot
```

## Summary

TPRS are effective at reducing speeds, increasing driver attentiveness, and minimizing erratic driving in work zones. Displacement under traffic remains a practical concern, especially for non-adhesive products, and more research is needed on long-term performance and optimal deployment strategiesState DOT’s increasingly use TPRS to alert drivers to upcoming work zones. Specifications for installing and deploying TPRS vary widely between states. Existing research overwhelmingly suggests that TPRS effectively gain drivers attention, improving speed limit compliance and reducing crash rates. The magnitude of these effects, however, vary widely among the studies considered in this review. This variance in effect may be a result of the variance in deployment practice, particularly in the spacing of strips in a TPRS array. The spacing of TPRS may also affect the rate of strip displacement, and the associated need for workers to enter the roadway. 
More research is needed to observe the actual rate of displacement of TPRS under high-volume, high-speed traffic in a real-world setting. No research identified in this review has addressed the spacing between strips within an array and how such spacing affects driver behavior in terms of speed reduction, braking behavior, or driver avoidance. Additionally, the research considered in this review did not generally observe driver behavior on mountainous, rural roads with high traffic speeds of the kind prevalent in Utah. These observations shaped the design of an experiment that will be the focus of the remainder of this report.
.