## Results and Analysis


```{r setup, file = "R/chapter_start.R", include = FALSE}
# a number of commands need to run at the beginning of each chapter. This
# includes loading libraries that I always use, as well as options for 
# displaying numbers and text.
library(modelsummary)
```

### Speeds
Analysis question 1: does having rumble strips slow down traffic?

```{r wavetronix}
# join observation information to the wavetonix data
tar_load(wavetronix)
tar_load(observations)

wavetronix <- wavetronix |>
  left_join(observations, by = c("site", "date")) 

```

```{r wgrp}
wavetronix |>
  group_by(site, strip_spacing, unit) |>
  summarise(speed_85 = mean(speed_85))
```

This is a t-test of means for one road comparing 35 ft spacing with no rumble strips.
```{r wtt}
groups <- wavetronix |>
  group_by(site, unit, strip_spacing) |> 
  select(speed_85) |>
  nest(-site, -unit, -strip_spacing) 

t.test(groups$data[[5]], groups$data[[7]])
t.test(groups$data[[9]], groups$data[[10]])
```




```{r wlm}
# estimate a model of the effect of rumble strips on speed
wlms <- wavetronix |>
  group_by(site) |>
  nest(data = -site) |>
  mutate(
    model = map(data, ~lm(speed_85 ~ unit, data = .x))
  )

list <- wlms$model |> setNames(wlms$site)
```


```{r}
modelsummary(list,  stars = TRUE)
```


## Braking and Avoidance

```{r}
#| label: tbl-avoidglm
#| tbl-cap: Binomial Models of TPRS Avoidance
tar_load(avoid_models)
modelsummary(avoid_models,  stars = TRUE)
```

```{r}
#| label: tbl-brakeglm
#| tbl-cap: Multinomial Models of TPRS Braking
tar_load(brake_models)
modelsummary(brake_models, shape = term ~ response,  stars = TRUE)
```



### Strip Displacement

```{r}
#| label: fig-momentum
#| fig-cap: Impact momentum for each transition, colored by site.
tar_load(momentum_per_transition_site)

momentum_per_transition_site
```

@fig-momentum shows the impact momentum for each transition, colored by site. The impact momentum is the change in speed for a vehicle in a given transition. The change in speed is calculated by subtracting the speed at the beginning of the transition from the speed at the end of the transition. This is a useful metric for understanding the effect of rumble strips on traffic flow.

### Worker Exposure

```{r cdf-sites}

tar_load(cdf_plots)

cdf_plots$site_plot
```