# Results and Analysis {#sec-results}

This section presents the results of the methodology in @sec-methodology.


```{r setup, include = FALSE}
# a number of commands need to run at the beginning of each chapter. This
# includes loading libraries that I always use, as well as options for 
# displaying numbers and text.
source("R/chapter_start.R")

library(modelsummary)
library(scales)

signed_log10 <- trans_new(
  name = "signed_log10",
  transform = function(x) sign(x) * log10(abs(x)),
  inverse = function(x) sign(x) * 10^abs(x)
)

spacing_palette <- c(
  "NO TPRS" = "#2E4756",
  "UDOT" = "#3C7A89",
  "1:2" = "#F49D37",
  "LONG" = "#D81159"
)

site_palette <- c(
  "US-6" = "#3C7A89",
  "US-191" = "#F49D37", 
  "SR-12" = "#D81159",
  "I-70" = "#2E4756"
)

cm <- c(
  "classTruck" = "Trucks",
  "classMotorcycle" = "Motorcycles",
  "spacing_typeUDOT" = "UDOT Spacing",
  "spacing_type1:2" = "1:2 Spacing",
  "spacing_typeLONG" = "Long Spacing",
  "siteSR-12" = "SR-12",
  "siteUS-191" = "US-191",
  "siteUS-6" = "US-6"
)

library(ggplot2)
theme_set(
  theme_minimal(base_family = "Times New Roman", base_size = 12)
)
```

## Speeds

@fig-speed shows the change in speed between units, 95% confidence bound of different between 85th percentile speeds. The 95% confidence bound is the range of speeds that are likely to occur if the null hypothesis is true. The confidence bounds are calculated using the bootstrap method. The 95% confidence bound is calculated by taking the upper and lower bounds of the 95% confidence interval, and then taking the difference between them. This is a useful metric for understanding the effect of rumble strips on traffic flow.


```{r}
#| label: fig-speed
#| fig-cap: Change in speed between units, 95% confidence bound of different between 85th percentile speeds.
tar_load(confidence_bounds)

# see plot_confidence_bounds() in R/speed.R for how this plot is made
confidence_bounds +
  scale_color_manual(values = spacing_palette) +
  labs(
    x = "Speed Difference [mph]",
    y = "Site",
    color = "Spacing Type"
  )
```

## Braking Behavior

```{r eval=FALSE}
## NOT going to include this in the paper
#| label: fig-braking
#| fig-cap: Braking behavior by spacing of TPRS and vehicle type.
tar_load(braking_plot)
braking_plot
```

```{r eval=FALSE}
#| label: tbl-brakeglm
#| tbl-cap: Multinomial Models of TPRS Braking
tar_load(brake_models)
modelsummary(brake_models, shape = term ~ response,  stars = TRUE, coef_rename = cm)
```

## Avoidance Behavior

```{r}
#| label: fig-avoidance
#| fig-cap: Avoidance behavior by spacing of TPRS and vehicle type.
tar_load(departure_plot)
departure_plot
```

```{r}
#| label: tbl-avoidglm
#| tbl-cap: Binomial Models of TPRS Avoidance
tar_load(avoid_models)
modelsummary(avoid_models,  stars = TRUE, coef_rename = cm)
```


### Strip Displacement

```{r}
#| label: fig-momentum-site
#| fig-cap: Impact momentum for each transition, colored by site.
tar_load(momentum_per_transition_site)
momentum_per_transition_site
```

```{r}
#| label: fig-momentum-spacing
#| fig-cap: Impact momentum for each transition, colored by spacing.
tar_load(momentum_per_transition_spacing)
momentum_per_transition_spacing
```

@fig-momentum-site shows the impact momentum for each transition, colored by site. The impact momentum is the change in speed for a vehicle in a given transition. The change in speed is calculated by subtracting the speed at the beginning of the transition from the speed at the end of the transition. This is a useful metric for understanding the effect of rumble strips on traffic flow.

### Worker Exposure

```{r}
#| label: fig-cdf
#| fig-cap: CDF of headways for each site.
tar_load(headway_cdf_site_plot)
headway_cdf_site_plot +
    coord_cartesian(xlim = c(0, 100), ylim = c(0, 1.05)) +
    scale_color_manual(values = site_palette) +
    scale_y_continuous(labels = scales::percent) +
    labs(
      x = "Headway (s)",
      y = "Cumulative %",
      color = "Site"
    )
```