# Results and Analysis {#sec-results}

This section presents the results of the methodology in @sec-methodology.


```{r setup, file = "R/chapter_start.R", include = FALSE}
# a number of commands need to run at the beginning of each chapter. This
# includes loading libraries that I always use, as well as options for 
# displaying numbers and text.
library(modelsummary)
library(scales)

signed_log10 <- trans_new(
  name = "signed_log10",
  transform = function(x) sign(x) * log10(abs(x)),
  inverse = function(x) sign(x) * 10^abs(x)
)
```

## Speeds

@fig-speed shows the change in speed between units, 95% confidence bound of different between 85th percentile speeds. The 95% confidence bound is the range of speeds that are likely to occur if the null hypothesis is true. The confidence bounds are calculated using the bootstrap method. The 95% confidence bound is calculated by taking the upper and lower bounds of the 95% confidence interval, and then taking the difference between them. This is a useful metric for understanding the effect of rumble strips on traffic flow.


```{r}
#| label: fig-speed
#| fig-cap: Chnage in speed between units, 95% confidence bound of different between 85th percentile speeds.
tar_load(confidence_bounds)
confidence_bounds
```

## Braking and Avoidance

```{r}
#| label: tbl-avoidglm
#| tbl-cap: Binomial Models of TPRS Avoidance
tar_load(avoid_models)
modelsummary(avoid_models,  stars = TRUE)
```

```{r}
#| label: tbl-brakeglm
#| tbl-cap: Multinomial Models of TPRS Braking
tar_load(brake_models)
modelsummary(brake_models, shape = term ~ response,  stars = TRUE)
```




### Strip Displacement

```{r}
#| label: fig-momentum
#| fig-cap: Impact momentum for each transition, colored by site.
tar_load(momentum_per_transition_site)

momentum_per_transition_site
```

@fig-momentum shows the impact momentum for each transition, colored by site. The impact momentum is the change in speed for a vehicle in a given transition. The change in speed is calculated by subtracting the speed at the beginning of the transition from the speed at the end of the transition. This is a useful metric for understanding the effect of rumble strips on traffic flow.

### Worker Exposure

```{r cdf-sites}

tar_load(cdf_plots)

cdf_plots$site_plot
```