# Results {#sec-results}


```{r setup, include = FALSE}
# a number of commands need to run at the beginning of each chapter. This
# includes loading libraries that I always use, as well as options for 
# displaying numbers and text.
source("R/chapter_start.R")

library(modelsummary)
library(scales)

signed_log10 <- trans_new(
  name = "signed_log10",
  transform = function(x) sign(x) * log10(abs(x)),
  inverse = function(x) sign(x) * 10^abs(x)
)

spacing_palette <- c(
  "NO TPRS" = "#2E4756",
  "UDOT" = "#3C7A89",
  "1:2" = "#F49D37",
  "LONG" = "#D81159"
)

site_palette <- c(
  "US-6" = "#3C7A89",
  "US-191" = "#F49D37", 
  "SR-12" = "#D81159",
  "I-70" = "#2E4756"
)

cm <- c(
  "classTruck" = "Trucks",
  "classMotorcycle" = "Motorcycles",
  "spacing_typeUDOT" = "UDOT Spacing",
  "spacing_type1:2" = "1:2 Spacing",
  "spacing_typeLONG" = "Long Spacing",
  "siteSR-12" = "SR-12",
  "siteUS-191" = "US-191",
  "siteUS-6" = "US-6"
)

gm <- c()

library(ggplot2)
theme_set(theme_minimal())
```

## Speeds

The Wavetronix radar devices record the mean and 85th percentile speeds for
vehicles passing the detectors aggregated to 15-minute bins, meaning that in an
8-hour observation period, 48 85th-percentile speeds are recorded at each
detector.   
@fig-speed shows the difference in 85th percentile speeds between Wavetronix 1
and Wavetronix 2 and the 95% confidence interval of this difference. A positive difference implies that 
the 85th percentile speed *increased* between Wavetronix 1 and Wavetronix 2. 
The results of this analysis do not show any clear correlation between change in
speed and TPRS spacing, though in many conditions there is a reduction in speed. If the TPRS spacing had a consistent impact on speed,
the relative position between the spacing types would be consistent between
sites. The data do not show this pattern; rather, the impact of TPRS presence or
spacing varies inconsistently between sites. Notably, there are cases on US-191
where vehicle speeds increased after drivers traversed the array of TPRS. While
this might seem to contradict the current consensus found in the literature,
some studies did acknowledge that TPRS influence on speed is highly localized [@mcavoyEvaluationTemporaryRemovable2009; @shahinEffectsHighwayWork2023]

```{r}
#| label: fig-speed
#| fig-cap: Change in 85th percentile speed between radar speed units with 95% confidence interval on difference in speeds.
tar_load(confidence_bounds)

# see plot_confidence_bounds() in R/speed.R for how this plot is made
confidence_bounds +
  scale_color_manual(values = spacing_palette) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(
    x = "Speed Difference [mph]",
    y = "Site",
    color = "Spacing Type"
  )
```

## Braking Behavior

Using Camera 1, the research team tracked drivers’ braking response to TPRS as
another proxy measure of driver awareness. @tbl-brakeglm shows a multinomial
logit model of braking behavior independently estimated for each site. In this
model, significant positive numbers are correlated with an increase in the
associated behavior relative to the not-braking reference. Vehicle type is not
consistently and significantly correlated with braking behavior. Motorcycles and
trucks are both more likely than passenger cars to brake before the strips at
I-70, trucks more likely at US-191, and motorcycles at US-6. There is no
correlation between vehicle type and after-strip braking at any site.
TPRS spacing is also not consistently and significantly correlated with braking behavior. Relative to the no-TPRS condition, before-strip braking is more likely at SR-12, US-6, and I-70 at all spacing types. Before-strip braking at US-191 is oddly somewhat less likely with strips installed than without strips in place. After-strip braking is more likely with strips installed on I-70, and with UDOT spacing on US-6. But after-strip braking is less common with strips installed on SR-12 than without strips installed. 
The fact that many vehicles brake before TPRS may suggest that the influence of TPRS is primarily visual rather than tactile for most road users. If drivers are reacting to TPRS on sight rather than striking the strips, this could explain why motorcycle drivers seemed to brake more when TPRS are not present. Motorcycle drivers may be braking before they enter the camera’s field of view when they see the TPRS installed, and thus not being counted. But when TPRS are not installed, motorcycle drivers may be braking later in response to seeing the lane closure rather than the TPRS. This interpretation is speculative and more research is needed. 

```{r eval=FALSE}
## NOT going to include this in the paper
#| label: fig-braking
#| fig-cap: Braking behavior by spacing of TPRS and vehicle type.
tar_load(braking_plot)
braking_plot
```

```{r }
#| label: tbl-brakeglm
#| tbl-cap: Multinomial Models of TPRS Braking
tar_load(brake_models)

mcfadden_rho2 <- function(model) {
  observed <- model$residuals + model$fitted.values
  counts <- colSums(observed)
  counts <- counts[counts > 0]
  null_loglik <- sum(counts * log(counts / sum(counts)))

  data.frame(
    mcfadden_rho2 = 1 - as.numeric(logLik(model)) / null_loglik
  )
}

brake_gof <- data.frame(
  raw = c("nobs", "rmse", "mcfadden_rho2"),
  clean = c(
    "Num.Obs.",
    "RMSE",
    if (knitr::is_latex_output()) {
      "McFadden $\\rho^2$"
    } else {
      "McFadden &rho;<sup>2</sup>"
    }
  ),
  fmt = c(0, 3, 3)
)

modelsummary(
  brake_models[-1],
  shape = term ~ response,
  stars = TRUE,
  coef_rename = cm,
  gof_function = mcfadden_rho2,
  gof_map = brake_gof,
  escape = FALSE
)
```

## Avoidance Behavior

The research team observed lane departures where vehicles appeared to avoid TPRS. It is the opinion of the researchers who watched the camera footage that most vehicles that departed their lane to avoid TPRS did so in an orderly manner similar to passing another vehicle, as opposed to a chaotic adjustment. There was no differentiation in the observations between partial lane departures, where only one side of tires crossed the center line or right-side line, and full lane departures, where all tires crossed either edge of the lane. On I-70, where TPRS were laid across both lanes, there were approximately 10 vehicles that drove fully in the shoulder to avoid the TPRS, but these were not classified separately. 

@tbl-avoidglm presents a series of binomial logit models estimating the probability of lane departure at each site. In this model, significant positive coefficients indicate that vehicles are more likely to avoid TPRS. Motorcycles have a consistent and dramatically higher rate of lane departure than either passenger vehicles or trucks, but trucks and passenger vehicles avoid at the same rate. This finding confirms prior literature that motorcycles, as vulnerable road users and more agile vehicles, are both motivated and capable of avoiding the TPRS. While some passenger vehicles and trucks do leave their lane to avoid the TPRS, the vast majority do not. 
The significance of the model coefficients showing TPRS avoidance at most TPRS layouts is a feature of TPRS avoidance being impossible when they are not deployed, which is the reference alternative in the model, and not a finding of importance. What is more important is that the estimated coefficients associated with TPRS spacing are not significantly different from each other at each site, indicating that the spacing does not incite or discourage TPRS avoidance.

```{r eval=FALSE}
#| label: fig-avoidance
#| fig-cap: Avoidance behavior by spacing of TPRS and vehicle type.
tar_load(departure_plot)
departure_plot
```

```{r}
#| label: tbl-avoidglm
#| tbl-cap: Binomial Models of TPRS Avoidance
tar_load(avoid_models)
avoid_gof <- data.frame(
  raw = c("nobs", "rmse", "logLik"),
  clean = c(
    "Num.Obs.",
    "RMSE",
    "Log Likelihood"
  ),
  fmt = c(0, 3, 3)
)
modelsummary(avoid_models[-1],  stars = TRUE, coef_rename = cm, gof_map = avoid_gof)
```


### Strip Displacement

```{r}
#| label: fig-momentum-site
#| fig-cap: Impact momentum for each transition, colored by site.
tar_load(momentum_per_transition_site)
momentum_per_transition_site
```

```{r}
#| label: fig-momentum-spacing
#| fig-cap: Impact momentum for each transition, colored by spacing.
tar_load(momentum_per_transition_spacing)
momentum_per_transition_spacing
```

@fig-momentum-site shows the impact momentum for each transition, colored by site. The impact momentum is the change in speed for a vehicle in a given transition. The change in speed is calculated by subtracting the speed at the beginning of the transition from the speed at the end of the transition. This is a useful metric for understanding the effect of rumble strips on traffic flow.

### Worker Exposure

```{r}
#| label: fig-cdf
#| fig-cap: CDF of headways for each site.
tar_load(headway_cdf_site_plot)
headway_cdf_site_plot +
    coord_cartesian(xlim = c(0, 60), ylim = c(0, 1.05)) +
    scale_color_manual(values = site_palette) +
    scale_y_continuous(labels = scales::percent) +
    labs(
      x = "Headway (s)",
      y = "Cumulative %",
      color = "Site"
    ) + theme_bw()
```
